---
title       : Statistical Methods with R
subtitle    : 
author      : Johnson Hsieh (謝宗震)
job         : DSP團訓班2
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow       # 
license     : 
logo        : 
widgets     : [mathjax, bootstrap, quiz]
mode        : selfcontained # {standalone, draft}


---

```{r setup, include=FALSE,echo = F, message = F, warning = F, tidy = F}
# make this an external chunk that can be included in any file
options(width = 100,digits=3,warn=FALSE)
opts_chunk$set(message = FALSE, eval=TRUE,fig.align = "center", warning = FALSE, comment = NA, dpi = 100, fig.width=6, fig.height=4,tidy = FALSE, cache = FALSE)

options(xtable.type = 'html')
knit_hooks$set(inline = function(x) {
  if(is.numeric(x)) {
    round(x, getOption('digits'))
  } else {
    paste(as.character(x), collapse = ', ')
  }
})
knit_hooks$set(plot = knitr:::hook_plot_html)
```

## Links To Slides
### Link to all slides
http://JohnsonHsieh.github.com/dsp-statR

### Source code on github
http://github.com/JohnsonHsieh/dsp-statR/

--- .segue .dark
## Intro R basic (conti.)

---
## 基礎教學
```{r}
# dat <- read.csv("http://johnsonhsieh.github.io/dsp-introR/data/hsb.csv")
dat <- read.csv("data/hsb.csv")
str(dat) # show dataset structure
```

---
## Homework
```{r eval=FALSE}
# Q:男女生考試總分的平均成績誰比較高，高多少分
# Hint:
score <- dat[, c("read","write","math","science","socst")]
mean(dat[dat$sex=="female", "math"])
mean(dat[dat$sex=="male", "math"])
sum()
```

---
## Homework
```{r}
# Q:男女生考試總分的平均成績誰比較高，高多少分
# A1:
score <- dat[, c("read","write","math","science","socst")]
F <- sum(score[dat$sex=="female",])/sum(dat$sex=="female")
M <- sum(score[dat$sex=="male",])/sum(dat$sex=="male")
F - M
```


---
## Homework
```{r}
# Q:男女生考試總分的平均成績誰比較高，高多少分
# A2:
score <- dat[, c("read","write","math","science","socst")]
total <- score[,1]+score[,2]+score[,3]+score[,4]+score[,5]
F <- mean(total[dat$sex=="female"])
M <- mean(total[dat$sex=="male"])
F - M
```

---
## Labeling Categorical Vars
```{r}
?factor
race <- factor(dat$race, 
               levels = c("Hispanic", "Asian", "African American", "White"), 
               labels = c("HI", "AS", "AA", "WH"))
head(race, n=10)
race2 <- factor(dat$race, 
               levels = c("African American", "White"), 
               labels = c("B", "W"))
head(race2, n=10)
```

--- 
## Sorting Data
```{r}
dat <- dat[order(dat$id, dat$sex), ] # sort data by id and sex
head(dat)
```

---
## Merging Data
- Use rbind or cbind
```{r}
dat$score <- rowSums(dat[,7:11]) # read + write + math + science + socst
# rowSums, colSums, rowMeans, colMeans
str(dat$score)
dat.female <- dat[dat$sex=="female",]
dat.male <- dat[dat$sex=="male",]
dat.both <- rbind(dat.female, dat.male)
cbind(dim(dat.female), dim(dat.male),dim(dat.both))
```

--- .segue .dark
## 台北市實價登錄資料

---
## Read data
```{r}
f <- file("data/dsp-gift-2013-big5/買賣st_A_10109_10109.csv", encoding="big5")
tab <- read.csv(f)
names(tab)
```

---
## 台北市實價登錄資料
```{r}
str(tab)
```

---
## Modifying and Managing Data
```{r}
tab1 <- tab # 整理資料
tab1$行政區 <- tab1$鄉鎮市區
tab1$車位 <- factor(tab1$交易標的, levels=c("房地(土地+建物)", "房地(土地+建物)+車位"),
                    labels=c("無", "有"))
tab1$型態 <- factor(tab1$建物型態, levels=c("公寓(5樓含以下無電梯)","華廈(10層含以下有電梯)",
                    "住宅大樓(11層含以上有電梯)"), labels=c("公寓","華廈","大樓"))
tab1$房價 <- (tab1$總價.元 - tab1$車位總價.元.)
?sub 
tab1$建物移轉總面積.平方公尺. <- as.numeric(sub("\\,", "",
                                          as.character(tab1$建物移轉總面積.平方公尺.))) 
tab1$面積 <- 0.3025*(tab1$建物移轉總面積.平方公尺. - tab1$車位移轉總面積.平方公尺.)
tab1$管理 <- tab1$有無管理組織
tab1$屋齡 <- 103 - tab1$建築完成年月%/%10000 #商數
tab1 <- tab1[, c("行政區","型態","房價","面積","屋齡","車位","管理")]
```

---
## Modifying and Managing Data
```{r}
head(tab1)
```

---
## Modifying and Managing Data
```{r}
summary(tab1)
tab1 <- na.omit(tab1)
tab1 <- tab1[tab1$房價!=0,]
```

--- .segue .dark
## Exploratory Data Analysis

---
## Tabulating Data
```{r}
?table
table(tab1[,1])
table(tab1[,2])
```

---
## Pivot tables
```{r}
# install.packages("reshape")
library(reshape)
?cast
cast(tab1, 型態 ~ ., fun.aggregate=mean, value="房價")
```
---
## Pivot tables
```{r}
cast(tab1, 行政區 ~ ., fun.aggregate=table, value="型態")
# cast(tab1, 行政區 ~ 管理, fun.aggregate=table, value="型態")
```

---
## 小挑戰
- 請計算台北市各行政區為的平均房價
- 請計算台北市各行政區各種建物型態的房價中位數
- Hint: use cast() function in reshape package, 中位數函數 median()

---
## Bar Chart and Pie Chart
- Ref: http://www.statmethods.net/graphs/, Quick-R
```{r, fig.width=8}
par(family="STHeiti", mfrow=c(1,2)) # Mac 中文字型設定
barplot(sort(table(tab1$行政區), decreasing=TRUE), las=2)
pie(sort(table(tab1$型態), decreasing=TRUE))
```

---
## Histogram and denstiy
- http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf, colors in R
```{r, fig.width=8, fig.height=3.5}
par(family="STHeiti", mfrow=c(1,2)) # Mac 中文字型設定
hist(tab1$面積)
hist(tab1$面積, breaks=10, col="lightblue", prob=TRUE, xlab="面積", main="Histogram")
lines(density(tab1$面積, bw=8), col=2, lwd=2)
```

---
## 小挑戰
- ?hist, ?density
- 改變 hist() 函數中的breaks參數，觀察直方圖的變化
- 改變 density() 函數中的bw參數，觀察機率密度函數的變化

---
```{r, fig.width=8, fig.height=4.5}
par(family="STHeiti", mfrow=c(2,2)) # Mac 中文字型設定
hist(tab1$房價, col="gray")
hist(tab1$面積, col="gold")
hist(log10(tab1$房價), col="darkgreen")
hist(log10(tab1$面積), col="lightgreen")
```

---
## Scatterplot
```{r, fig.width=8, fig.height=4}
par(family="STHeiti", mfrow=c(1,2)) # Mac 中文字型設定
plot(tab1$房價/10^4, tab1$面積, xlab="房價(萬元)", ylab="面積(坪)")
plot(tab1$房價/10^4, tab1$面積, xlab="log房價(萬元)", ylab="log面積(坪)", log="xy")
```

---
## Line chart
```{r}
attach(beaver1)
hours <- time %/% 100 + 24*(day - day[1]) + (time %% 100)/60
plot(hours, temp, type="l", main="Body Temperature Series")
```

---
## Correlation
```{r}
x <- tab1$房價/10^4
y <- tab1$面積
cor(x, y)
cor(log10(x), log10(y))
```

---
## Regression
```{r, fig.width=6, fig.height=4}
par(family="STHeiti") # Mac 中文字型設定
plot(x, y, xlab="房價(萬元)", ylab="面積(坪)", log="xy")
fit <- lm(log10(y) ~ log10(x)) # log10.y = -0.85 + 0.73 * log10.x
abline(fit, col=2, lwd=2)
```

---
```{r}
summary(fit) 
```

---
## Prediction
```{r}
new <- data.frame(x = c(500, 800, 1600, 2500, 5000, 8000))
10^predict(fit, newdata=new)
```

---
## Multiple Regression
```{r}
fit1 <- lm(log10(面積)~log10(房價/10^4),data=tab1)
fit2 <- lm(log10(面積)~log10(房價/10^4)+型態+車位+管理,data=tab1)
fit2
```

---
```{r}
summary(fit1)
```

---
```{r}
summary(fit2)
```

---
```{r}
anova(fit1, fit2)
```

---
```{r}
par(family="STHeiti") # Mac 中文字型設定
cols <- adjustcolor(ifelse(tab1$型態=="公寓", 2, 4), 0.3)
plot(面積 ~ 房價, data=tab1,  col = cols, pch=19, log="xy")
legend("topleft", c("公寓", "華廈+大樓"), pch=19, col=c(2,4), bty="n")
```


---
## References
- [Introducing R](http://www.ats.ucla.edu/stat/r/seminars/intro.htm), UCLA R seminars
- [R的資料型態](http://rpubs.com/wush978/R_DataType), TW R User Group
- [Quick R graphs](http://www.statmethods.net/graphs/), Quick R
- [免費電子書 -- R 統計軟體](http://ccckmit.wikidot.com/r:main), 陳鍾誠的網站
- [StackOverflow](http://stackoverflow.com/), getting help online
